On the Functionality of the Multilayer Perceptron

### 2020-09-13 Thesis/Sections/TheNerualNetwork.tex
12 Sept 2020
- Updates to Feature Selections

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[top=2.5cm,left=2.5cm,right=2.5cm]{geometry}

@@

\subsection{An Introduction to Neural Networks}

@@

\subsection{The Structure of a Neural Network}

@@

\subsection{Layers Used in Classification Neural Network}

@@

\paragraph*{}Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer for short, was one of the earliest function types used in artificial neural network models. A dense layer is simply composed of a layer of \textit{artifical neurons}, each of which holds a numerical value within it, called the \textit{activation} of that neuron. This idea was developed back from McCulloch and Pitts' work \cite{McCulloch}, and was expanded upon by Frank Rosenblatt in  1957 \cite{Geron}.

@@

\paragraph*{}Suppose the layer $l-1$ contains $m$ neurons. Then the weighting-matrix, $W^{(l)}$ has shape $m \times n$, the bias-vector $b^{(l)}$ has shape $m \times 1$.
Thus for a dense layer $l$, the exact values of each activation is given by \cite{Geron,Loy}
\begin{equation}
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{n-1}
\end{bmatrix}^{(l)} =
\sigma^{(l)} \Bigg(
\begin{bmatrix}
w_{0,0} & w_{0,1} & \hdots & w_{0,m-1} \\
w_{1,0} & w_{1,1} & \hdots & w_{1,m-1} \\
\vdots & \vdots & \ddots & \vdots  \\
w_{n-1,0} & w_{n-1,1} & \hdots & w_{n-1,m-1} 
\end{bmatrix}^{(l)} 
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{m-1}
\end{bmatrix}^{(l-1)}
\begin{bmatrix}
b_0 \\ b_1 \\ \vdots \\ b_{n-1}
\end{bmatrix}^{(l)} \Bigg)
\end{equation}
or more compactly:
\begin{equation}
\label{eqn-DenseFeedForward}
x^{(l)} = \sigma^{(l)} \Big( W^{(l)} x^{(l-1)} + b^{(l)} \Big)
\end{equation}
Eqn. (\ref{eqn-DenseFeedForward}) is generally referred to as the \textit{dense layer feed-forward equation} \cite{Goodfellow}.

@@

@@

@@

@@

\subsection{Training a Neural Network}

\paragraph*{}A neural network's job is to produce an approximation of some function $F$, which we call $F^*$. The model must have a procedure for updating the parameters in $\Theta$ to allow for a reasonable approximation of $F$. To better understand this, we turn to Tom Mitchell's explanation of a learning algorithm \cite{Goodfellow,Mitchell}:
\begin{quote}
A computer program is said to learn from experience $E$ with respect to some class of tasks $T$ and performance measure $P$, if it's performance at tasks in $T$, as measured by $P$, improves with experience $E$.
\end{quote}
Without any external intervention, a model must update itself to improve it's performance at a give task a new information is presented to it. To do this, the model must be constructed with a training procedure in mind.

\paragraph*{}If we consider the set of parameters $\theta$ as a the concatenation, of each layer's weighting matrix and bias vector such that:
\begin{equation}
\label{eqn-Theta}
\theta = \big\{
W^{(0)}, b^{(0)}, W^{(1)}, b^{(1)}, W^{(2)}, b^{(2)}, ... ,W^{(L-2)}, b^{(L-2)},W^{(L)}, b^{(L-1)},
\big\}
\end{equation}
We can consider each parameter is simply a float-point number. Each value for each number contribute to the output of the function. In increasing complicated neural networks, there can be upwards of hundreds of thousands,or even millions of elements in $\theta$ making a neural network a functions that exist is  a very high dimensional parameter-space.

\paragraph*{}Finally,

@@

\paragraph*{}The success of a neural network algorithm is enormously dependent of the strength of the chosen features, rather than the quantity of available data \cite{Geron,Khan,Liu}. We have derived an appropriate set of inputs arrays \textit{[See Section on Features]}, and constructed a neural network architecture to compliment these features.

@@

\paragraph*{}The spectrogram branch is pictured on the left side of fig. (\ref{fig-NetworkArchitecture}). A spectrogram is a representation of a sound wave, or signal by representing energy distributions as a function of \textit{time} and \textit{frequency} \cite{White,Olsen,Kahn}. \textcolor{blue}{Details on the creating of the spectrogram can be found in the \textit{features} section of this document.}

@@

\paragraph*{}The input layer of the perceptron accepts a 2-Dimensional array. The axes, in order of indexing, represent (i) the size of the \textit{mini-batch} of samples, (ii) the number of features for each sample. We use the same model hyper-paramater of $64$ samples per batch, and have developed $23$ unique clssification features that have been derived from time-space and frequency-space representations of the audio file data. We can denote the 2D shape if the input object into this branch as:
\begin{equation}
\label{eqn-shapeX2}
X_2 \in \mathbb{R}^{(64 \times 23)}
\end{equation}
Again, any other shape will be rejected by the model, and an error is raised.

>>>>

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[top=2.5cm,left=2.5cm,right=2.5cm]{geometry}

@@

\subsection{An Introduction to Neural Networks}
\label{sec-NerualNetworkIntro}

@@

\subsection{The Structure of a Neural Network}
\label{sec-NetworkStructre}

@@

\subsection{Layers Used in Classification Neural Network}
\label{sec-Layers}

@@

\paragraph*{}The combination of eqn. (\ref{eqn-matVectEqn}) and eqn. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in eqn. (\ref{eqn-altLayerFunction}). Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer for short, was one of the earliest function types used in artificial neural network models \cite{Geron2,Loy}. They are among the most commonly used layer types. A dense layer is composed of a layer of \textit{artifical neurons}, each of which holds a numerical value within it, called the \textit{activation} of that neuron. This idea was developed back from McCulloch and Pitts' work \cite{McCulloch}, and was expanded upon by Frank Rosenblatt in  1957 \cite{Geron}.

@@

\paragraph*{}Suppose the layer $l-1$ contains $m$ neurons. Then the weighting-matrix, $W^{(l)}$ has shape $m \times n$, the bias-vector $b^{(l)}$ has shape $m \times 1$. We can denote the function in a similar manner to eqn.(\ref{eqn-altLayerFunction}) as:
\begin{equation}
\label{eqn-FunctionDense}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(m \times 1)} \rightarrow x^{(l)} \in \mathbb{R}^{(n \times 1)}
\end{equation}

Thus for a dense layer $l$, the exact values of each activation is given by \cite{Geron,Loy}:
\begin{equation}
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{n-1}
\end{bmatrix}^{(l)} =
\sigma^{(l)} \Bigg(
\begin{bmatrix}
w_{0,0} & w_{0,1} & \hdots & w_{0,m-1} \\
w_{1,0} & w_{1,1} & \hdots & w_{1,m-1} \\
\vdots & \vdots & \ddots & \vdots  \\
w_{n-1,0} & w_{n-1,1} & \hdots & w_{n-1,m-1} 
\end{bmatrix}^{(l)} 
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{m-1}
\end{bmatrix}^{(l-1)} +
\begin{bmatrix}
b_0 \\ b_1 \\ \vdots \\ b_{n-1}
\end{bmatrix}^{(l)} \Bigg)
\end{equation}
or more compactly:
\begin{equation}
\label{eqn-DenseFeedForward}
x^{(l)} = \sigma^{(l)} \Big( W^{(l)} x^{(l-1)} + b^{(l)} \Big)
\end{equation}
Eqn. (\ref{eqn-DenseFeedForward}) is generally referred to as the \textit{dense layer feed-forward equation} \cite{Goodfellow}.

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}). This layer get's it's name from it's implementation of the mathematical \textit{convolution} function. The 2D discrete-convolution of function $A(i,j)$ and $B(i,j)$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C(i,j) = (A * B)(i,j) = \sum_{u}\sum_{v} A(i,j)B(i - u,j - v)
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)})$, by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(N \times M)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activation of the $x^{(l-1)}$ array. The step size in each of the 2-dimension is known as the \textit{stride}. Each of the $K$ filters is used to generate a $m \times n$ feature map from the input activation. For an appropriately trained network, some filters may be dedicated to detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Loy,Geron}.

\paragraph*{}In general, for an $N \times M$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(N \times M)} \rightarrow
x^{(l)} \in \mathbb{R}^{([N-n+1] \times [M-m+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(4 \times 3)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\label{fig-2DConvExample}
\begin{center}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
\caption{Input Activations}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
\caption{Convolution Filter}
\end{subfigure}
\begin{subfigure}{0.9\textwidth}
\centering
\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
\caption{Convolution Result}
\end{subfigure}
\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

\paragraph*{}For 2D convolution over input $x^{(l-1)}$ with feature map $W^{(l)}_k$, producing activations $x^{(l)}$, we compute the activations $x^{(l)}_{k}$ as \cite{Goodfellow}
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k = \sigma^{(l)}\bigg[ \Big( \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} x^{(l-1)}(i,j) W^{(l)}_k(i - u,j - v) \Big) + b^{(l)}_k \bigg]
\end{equation}
\textcolor{red}{Please check my math here!}
This operation repeats for each of the $K$ feature maps. Each maps has it's own $n \times m$ weighting matrix and appropriately shaped bias .

\paragraph*{}The convolution layers allows for several advantages. Among these are (i) \textit{sparse-connectivity}: not every single activation (pixel) is connect to every single output pixel, so it is more computationally efficient, (ii) \textit{positional invariance}: key features can be identified regardless of where they are in the layer

@@

\paragraph*{}A Maximum Pooling layers simply returns the maximum neuron activation in a group of neurons. In the case of 2D Max Pooling, we choose a kernel size to be $m \times n$, just like in the 2D Convolution layer, and extract the maximum value in each window, and then step along accrdoing to the chosen stride size \cite{Loy,Goodfellow}.As an example, consider again figure (\ref{fig-2DConvExample}). Using the $2 \times 2$ kernel on the $3 \times 4$ input, each box would then contain the maximum value of the input activations. We detail this in fig. (\ref{fig-2DMaxPool}):
\begin{figure}[h]
\label{fig-2DMaxPool}
\begin{center}
\includegraphics[scale=0.6]{../figures/2DConvExample(d)}
\end{center}
\caption{The result of 2D maximum-pooling an input array. This image was adapted from Loy, pg. 126 \cite{Loy}}
\end{figure}

\paragraph*{}Pooling layers, such as maximum pooling, average pooling, or similar are typically placed after a layer or a set of layers of convolution. The purpose of this arrangement is to reduce the number of weights, thereby dropping the complexity of the model and ensuring only features with large activation values are preserved \cite{Geron,Loy,Goodfellow}.

@@

\subsection{Activation Functions Used in Network Layers}
\label{sec-ActivationFunctions}

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{sec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in eqn. (\ref{eqn-matVectEqn}), and then an element-wise activation function as in eqn. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in out classification neural network.

\subsubsection{Rectified Linear Unit}

\subsubsection{Softmax}

@@

\subsection{Training a Neural Network}
\label{sec-Training}

\paragraph*{}A neural network's job is to produce an approximation of some function $F$, which we call $F^*$. The model must have a procedure for updating the parameters in $\Theta$ to allow for a reasonable approximation of $F$. To better understand this, we turn to Tom Mitchell's explanation of a learning algorithm \cite{Goodfellow,Mitchell}:
\begin{quote}
A computer program is said to learn from experience $E$ with respect to some class of tasks $T$ and performance measure $P$, if it's performance at tasks in $T$, as measured by $P$, improves with experience $E$.
\end{quote}
Without any human intervention, a model must update itself to improve it's performance at a give task a new information is presented to it. To do this, the model must be constructed with a training procedure in mind.

\paragraph*{}We consider the set of parameters $\Theta$ as a concatenation of each applicable layer's weighting matrix and bias vector such that:
\begin{equation}
\label{eqn-Theta}
\theta = \big\{
W^{(0)}, b^{(0)}, W^{(1)}, b^{(1)}, W^{(2)}, b^{(2)}, ... ,W^{(L-2)}, b^{(L-2)},W^{(L)}, b^{(L-1)},
\big\}
\end{equation}
We can consider each parameter as just float-point number, which is stored in a weight matrix $W^{(l)}$ or bias vector $b^{(l)}$. Each value for each parameter contributes to the output of the function. In increasingly complicated neural networks, there can be upwards of hundreds of thousands,or even millions of elements in $\theta$ making a neural network a functions that exist is a very high dimensional parameter-space.

\subsubsection{The Cost Function}

\paragraph*{}Suppose we pass a training sample $x^{0})$, with expected outcome $y$, into the neural network . For a perfectly accurate model, we expect the network's final layer, the output layer, $x^{(L-1)}$ to have activations such that $x^{(L-1)} \approx y$. Since the starting parameters of a network are often randomly generated, the prediction for training sample $x^(0)}$ is not likely to be similar to $y$ at all. We introduce a \textit{Cost Function} into our model that compares the difference between the neural output $x^{(L-1)}$ (also called $y^*$) and the desired, or expected output $y$ \cite{Goodfellow,James}.

\paragraph*{}There are various types of cost functions

\subsubsection{The Optimizer}

@@

\paragraph*{}The success of a neural network algorithm is enormously dependent of the strength of the chosen features, rather than the quantity of available data \cite{Geron,Kahn,Liu}. We have derived an appropriate set of inputs arrays \textit{[See Section on Features]}, and constructed a neural network architecture to compliment these features.

@@

\paragraph*{}The spectrogram branch is pictured on the right side of fig. (\ref{fig-NetworkArchitecture}). A spectrogram is a representation of a sound wave, or signal by representing energy distributions as a function of \textit{time} and \textit{frequency} \cite{White,Olsen,Kahn}. \textcolor{blue}{Details on the creating of the spectrogram can be found in the \textit{features} section of this document.}

@@

\paragraph*{}The input layer of the perceptron accepts a 2-Dimensional array. The axes, in order of indexing, represent (i) the size of the \textit{mini-batch} of samples, (ii) the number of features for each sample. We use the same model hyper-paramater of $64$ samples per batch, and have developed $23$ unique clssification features that have been derived from time-space and frequency-space representations of the audio file data. We can denote the 2D shape if the input object into this branch as:
\begin{equation}
\label{eqn-shapeX2}
X_2 \in \mathbb{R}^{(64 \times 20)}
\end{equation}
Again, any other shape will be rejected by the model, and an error is raised.
